---
title: "p8105_hw2_ym2813"
author: "Anna Ma"
date: "10/4/2021"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean the Mr.Trash Wheel sheet

```{r}
trash_wheel_df = 
  read_excel("data_hw2/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
             sheet = 1, col_names = TRUE, range = "A2:N534") %>% 
  janitor::clean_names() %>%  #clean the variable names
  drop_na(dumpster) %>% #omitting rows without dumpster-specific data
  rename(weight = weight_tons,volume = volume_cubic_yards) %>% #reasonable variable name 
  mutate(sports_balls = round(sports_balls,0)) #round sports_ball to integer

trash_wheel_df
```

Read and clean precipitation data for 2018 

```{r}
precipitation_2018_df = 
  read_excel("data_hw2/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
             sheet = 7, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2018") %>% 
  select(year,month,total)

precipitation_2018_df
```

Read and clean precipitation data for 2019

```{r}
precipitation_2019_df = 
  read_excel("data_hw2/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
             sheet = 6, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2019") %>% 
  select(year,month,total)

precipitation_2019_df
```

Combine precipitation datasets and convert month to a character variable

```{r}
precipitation_both_df = 
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>% 
  mutate(month = month.name[month])

precipitation_both_df
```

In the Mr. Trash Wheel data set, we have a total of `r nrow(trash_wheel_df)` observations, each representing a dumpster. The data has `r ncol(trash_wheel_df)` variables, collecting key information about Mr. Trash Wheel such as specific dumpster number, date, amount of trash they collected in weight and volume, types of trash they collected such as bottles, polystyrene, cigarette butts, and bags, and the homes they powered. The median number of sports balls in a dumpster in 2019 is `r trash_wheel_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`. 

In the data set for precipitation in 2018, we have a total of `r nrow(precipitation_2018_df)`observations, and `r ncol(precipitation_2018_df)` variables, which lists out month and precipitation within that month. 

Similarly, in the 2019 precipitation data set, we have a total of `r nrow(precipitation_2019_df)` observations and `r ncol(precipitation_2019_df)` variables that lists the month and  precipitation. 

In the combined data set, we have `r nrow(precipitation_both_df)` observations, and `r ncol(precipitation_both_df)` variables listing out the month with precipitations in both 2018 and 2019; and also the total amount of precipitation in each month and year. The total precipitation in 2018 is `r precipitation_both_df %>% filter(year == 2018) %>% pull(total) %>% sum()`

## Problem 2

Read and clean the data in pols_month

```{r, message=FALSE}
pols_month_df = 
  read_csv("data_hw2/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate("mon",c("year","month","day"),sep = "-") %>% 
  mutate(
    month = month.name[as.numeric(month)],
    president = case_when(prez_gop >= 1 ~ "gop",
                          prez_dem == 1 ~ "dem")
    ) %>% 
  select(-c(day,prez_dem,prez_gop))

pols_month_df
```

read and clean the data in snp.csv

```{r,message=FALSE}
snp_df = 
  read_csv("data_hw2/fivethirtyeight_datasets/snp.csv") %>%
  separate(date,c("month","day","year"),sep = "/") %>% 
  mutate(year = as.numeric(year),
         month = as.numeric(month),
         year = case_when(year <= 15 ~ 2000+ year ,
                          year > 15 ~ 1900+year)) %>% 
  arrange(year,month) %>% 
  mutate(
     month = month.name[month],
     year = as.character(year)) %>% 
  select(year,month,everything()) 

snp_df
```

Tidy the unemployment data set

```{r,message=FALSE}
unemployment_df = 
  read_csv("data_hw2/fivethirtyeight_datasets/unemployment.csv",
           col_names = c("year", "1":"12"),skip = 1) %>%
  pivot_longer(
    "1":"12",
    names_to = "month",
    values_to = "percentage") %>% 
  mutate(
     month = month.name[as.numeric(month)],
     year = as.character(year))

unemployment_df
```

Join the data sets

```{r,message=FALSE}
combined_df = 
  left_join(pols_month_df, snp_df, by = c("year","month")) %>% 
  left_join(unemployment_df, by = c("year","month"))

combined_df
```

The pols-month data set has `r nrow(pols_month_df)` observations and `r ncol(pols_month_df)` variables. It contained the number of national politicians who are democratic or republican in the 68 years ranging from January 1947 and June 2015. Some of the key variables includes the party of the president on the associated date, the number of republican or democratic governors, senators,or representatives on the associated date.

The snp data set has `r nrow(snp_df)` observations and `r ncol(snp_df)` variables. It contained the Standard & Poor's stock market index in 65 years ranging between January 3rd 1950 to July 1st 2015. The key variables are date of observation, including year, month, and day, and the closing values of the S&P stock index on the observation date. 

The unemployment data set has `r nrow(unemployment_df)` observations and `r ncol(unemployment_df)` variables. It contained the information about the percentage of unemployment of each month between 1948 to 2015, a range of 67 years. The key variables are year, month, and the percentage of unemployment. 

The resulting data set is composed of the pols-month, snp, and unemployment data set. It has `r nrow(combined_df)` observations and `r ncol(combined_df)` variables. This is because we merged snp and unemployment into the pols-month data set, which will keep everything in the pols-month and merge the snp and unemployment data by year and month.The final data set contained the number of national politicians who are democratic or republican, the Standard & Poor's stock market index, and the unemployment percentage of each month in 68 years, ranging from January 1947 and June 2015.

## Problem 3

Read and tidy the data

```{r,message=FALSE}
baby_names_df = 
  read_csv("data_hw2/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename("year" = year_of_birth, "name" = childs_first_name) %>%
  mutate(gender = str_to_lower(gender),
         name = str_to_lower(name),
         ethnicity = case_when(
         str_detect(ethnicity,"ASIAN") ~ "asian and pacific islander",
         str_detect(ethnicity,"BLACK") ~ "black non hispanic",
         ethnicity == "HISPANIC"  ~ "hispanic",
         str_detect(ethnicity,"WHITE") ~ "white non hispanic")) %>%
  distinct() # remove duplicate rows

baby_names_df
```

The data set has `r nrow(baby_names_df)` observations and `r ncol(baby_names_df)` variables, including birth year, gender, ethnicity, name, count, and rank. 

Table for the rank in popularity of the name “Olivia” as a female baby name over time

```{r}
olivia_df = baby_names_df %>% 
  filter(gender == "female", name == "olivia") %>% 
  select(-c(count,name,gender)) %>% 
  pivot_wider(
    names_from = "year",
    values_from = "rank")
knitr::kable(olivia_df, caption = "Rank of the Popularity of Olivia") 
```

Table showing the most popular name among male children over time.

```{r}
popular_male_name_df = baby_names_df %>% 
  filter(gender == "male", rank == 1) %>%
  select(-c(gender,rank,count)) %>% 
   pivot_wider(
    names_from = "year",
    values_from = "name"
  )
knitr::kable(popular_male_name_df, caption = "The Most Popular Name among Male Children Over Time") 
  
```

Scatter plot for the name popularity of male, white non-hispanic children born in 2016

```{r}
name_male_2016_df = baby_names_df %>% 
  filter(gender == 'male', ethnicity == 'white non hispanic', year == 2016) %>% 
  select(-c(year,gender,ethnicity))

ggplot(name_male_2016_df, aes(x = rank, y = count)) +
  geom_point() +
  ggtitle("Scatter Plot of Name Popularity for White non-hispanic Male Children Born in 2016") +
  ylab("Number of Children with A Name") + 
  xlab("Rank in popularity of the name")
```
